---
title: "Biological composition of points"
output: html_notebook
---

# Intro

Note to self: need to summarize by video quality in this script!

inputs:  (1) MarVid report (aka "species_abundance.csv") (in downloads folder) (2) Taxonary (on HI onedrive folder)
Workflow: edit the vme of interest
outputs: response data (aka "resp") as a simple dataframe. Includes the density of all vme indicators for the selected vme as well as total richness and total abundance.

# Libraries
```{r}
library(dplyr)
#install_github("trinker/qdapTools")
library(qdapTools)
library(readxl)
library(tidyr)
```

# Load data
```{r}
# for now use "fake marvid report" script

#data <- read.csv(file.path(Data, "species_abundance.csv"), sep = "|")
#data <- data %>% makesampid

taxonary <- read_xlsx(file.path(Data,"Taxonary.xlsx"), sheet = 1) # needs to be closed
```

# Lump detailed identifications with their corresponding taxon, and filter out generic/broad taxa
```{r}
# lump

# data <- data %>% mutate(taxon=lookup(taxon, data.frame(select(taxonary, c(Reference_List, Analysis_name))), missing=NULL)) %>%
#   group_by(taxon, SampID) %>%
#   summarize(density = sum(abundance.Nper100m.2)) %>%
#   select(SampID, taxon, density)
# 
# filter out generic taxa

head(data)
```

# Derive VME data
Example: Burgos et al classification

## available VME indicators
```{r}
print(unique(taxonary$VME_Burgos_etal_2020))
```
## Do they have the same environmental niche, approximately? (Jonatan)
Or are they within the same biotope, at least? Need to answer yes to one of these two questions before going ahead. If answer is no, then cut indicators out
```{r}
# set vme
vme <- "Hard bottom sponge aggregations"

```

## Filter data by VME indicators
```{r}
#derive response
resp1 <- data %>% group_by(SampID) %>%
  summarize(tot_rich = length(SampID), tot_dens = sum(density))
  
  
resp2 <- data %>% filter(taxon%in%taxonary$Reference_List[which(taxonary$VME_Burgos_etal_2020==vme)]) %>%
  group_by(SampID) %>%
  summarize(vmeind_density = sum(density))

resp <- left_join(resp1,resp2) %>% mutate(vmeind_density=replace_na(vmeind_density,0))
resp
```

## Aggregate to whole video line and calculate sd
```{r}
respw <- left_join(resp, sample_info) %>% group_by(video.line) %>%
  summarize(mean_rich = mean(tot_rich), mean_dens = mean(tot_dens), sd_dens = sd(tot_dens))
head(respw)
```

# Save output
```{r}
rm(resp1, resp2)
write.csv(resp, file.path(savepath, "resp.csv"))
```


