---
title: "Biological composition of points"
output: html_notebook
---

# Intro
Inputs:  (1) MarVid report (aka "species_abundance.csv") (in downloads folder, or use copy from Data folder)
(2) Taxonary (on HI onedrive folder, or use copy from Data folder)
Workflow: edit the vme of interest
Outputs: (1) response data (aka "resp") as a simple dataframe. Includes the density of all vme indicators for the selected vme as well as total richness and total abundance.

# Libraries
```{r}
library(dplyr)
#install_github("trinker/qdapTools")
library(qdapTools)
library(readxl)
library(tidyr)
```

# Load data
```{r}
# species densities (marvid data)
# for now use "fake marvid report" script

#data <- read.csv(file.path(Data, "species_abundance.csv"), sep = "|")
#data <- data %>% makesampid

# Taxonary
taxonary <- read_xlsx(file.path(Data,"Taxonary.xlsx"), sheet = 1) # needs to be closed
```

## Lump detailed identifications with their corresponding taxon, and filter out generic/broad taxa
Only applicable when using data directly downloaded from Marvid
```{r}
# lump

# data <- data %>% mutate(taxon=lookup(taxon, data.frame(select(taxonary, c(Reference_List, Analysis_name))), missing=NULL)) %>%
#   group_by(taxon, SampID) %>%
#   summarize(density = sum(abundance.Nper100m.2)) %>%
#   select(SampID, taxon, density)
# 
# filter out generic taxa

head(data)
```

# Derive VME data
Example: Burgos et al classification

## available VME indicators
```{r}
print(unique(taxonary$VME_Burgos_etal_2020))
```
## Do they have the same environmental niche, approximately? (Jonatan)
Or are they within the same biotope, at least? Need to answer yes to one of these two questions before going ahead. If answer is no, then cut indicators out
```{r}
# set vme
vme <- "Hard bottom sponge aggregations"

```

## Filter data by VME indicators
```{r}
#derive response
resp1 <- data %>% group_by(SampID) %>%
  summarize(tot_rich = length(SampID), tot_dens = sum(density))
  
  
resp2 <- data %>% filter(taxon%in%taxonary$Reference_List[which(taxonary$VME_Burgos_etal_2020==vme)]) %>%
  group_by(SampID) %>%
  summarize(vmeind_density = sum(density))

resp <- left_join(resp1,resp2) %>% mutate(vmeind_density=replace_na(vmeind_density,0))
resp
```

## Explore resp
```{r}
hist(resp$vmeind_density)
```

# Save output
```{r}
rm(resp1, resp2)
write.csv(resp, file.path(savepath, paste(vme,"resp.csv", collapse=" ")))
```


