---
author: "Havforskningsinstituttet"
date: "`r format(Sys.Date())`"
output: word_document
bibliography: references.bib
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
# libraries
library(ranger)
library(spm)
library(yardstick)
library(readxl)
library(dplyr)
library(tinytex)
library(knitr)
library(Ckmeans.1d.dp)
library(terra)
library(raster)
library(rgdal)

# paths
Data <- "G:/R_projects/VH_MarVid/Data"
outpath <- "G:/R_projects/VH_MarVid/Results"
savepath <- "G:/R_projects/VH_MarVid/Intermediate_objects"

# quantiles
quantiles = c((1-.682)/2, 0.5, 1-(1-.682)/2)

# taxonary
taxonary <- read_xlsx(file.path(Data,"Taxonary.xlsx"), sheet = 1) # needs to be closed

# vme and indicators
vme <- "Deep-sea sea pen communities"
remove="nothing"
#remove <- c("Mycale lingua", "Phakellia sp.", "Axinellidae", "Axinella infundibuliformis")
indicators <- taxonary$Reference_List[which(taxonary$VME_BuhlMortensen_etal_2023==vme)]
#indicators <- indicators[-which(indicators%in%remove)]

# model
name <-"m2"
m <- readRDS(file = file.path(outpath, paste(paste(vme,"minus", paste(substr(remove, 1,7), collapse=", ")),
                                             paste0(name, ".rda"))))

# re-generate v_test and v_train
# v <- read.csv(file.path(savepath, paste(vme, "minus", paste(substr(remove, 1,7), collapse=", "), "v.csv", collapse=" ")))
# v <- v %>% 
#       mutate_at(vars(  landscape, sedclass, seddan, sedmil    ), factor)
#data_split <- read.csv(file.path(outpath, paste(vme,"minus", paste(substr(remove, 1,7), collapse=", "), "data split.csv")))
#v_train <- v %>% left_join(filter(data_split, data_split$set=="train")) %>% filter(!is.na(set))
#v_test <- v %>% left_join(filter(data_split, data_split$set=="test")) %>% filter(!is.na(set))

# predict with m3
# vme.qrf2 <- predict(m,  v_test,
# type="quantiles", quantiles=quantiles)$predictions
# v_test$vme.qrf2pr = vme.qrf2[,2]
# v_test$vme.qrf2var = (vme.qrf2[,3]-vme.qrf2[,1])/2
# 
# # test m3
# predicted <- v_test$vme.qrf2pr
# observed <- v_test %>%
# dplyr::select(vmeind_density)
# vecv <- vecv(pull(observed),pull(data.frame(unlist(predicted))))
# ccc <- ccc_vec(pull(observed),predicted)

# threshold
distribution <- rast(file.path(file.path(outpath, name), paste(vme,"minus", paste(remove, collapse=", "), "pred.tif")))
error <- rast(file.path(file.path(outpath, name), paste(vme,"minus", paste(remove, collapse=", "), "pred var.tif")))


# data summary
sp<-indicators
e <- read.csv(file.path(savepath, "pred_data/e.csv"))
sample_info <- read.csv(file.path(Data,"sample_info.csv"))
sample_info <- sample_info %>% rename(video.line=VL, section=SampID2, mean.latitude=y_coord, mean.longitude=x_coord) %>%
  dplyr::select(video.line, section, mean.latitude, mean.longitude, SampID)
e <- cbind(e,sample_info)
data <- read.csv(file.path(Data,"species_densities.csv"))
data <- data %>% rename(taxon=clean_taxonomy, density = density_n100m2) %>% dplyr::select(SampID, taxon, density)
threshold <- vector(mode = "numeric",
                   length = length(sp))
for (i in 1:length(sp)){
  threshold[i] <-  quantile(sort(dplyr::pull(dplyr::filter(data, taxon==sp[i]))), c(.01, .05))[1]
}
thresholdt <- data.frame(sp=sp, threshold=as.numeric(threshold))
data_extract <- data %>% filter(taxon%in%indicators) %>% left_join(e, by = join_by(SampID == section))
data_extract_merged <-merge(data_extract,thresholdt,by.x="taxon",by.y="sp",all.x=TRUE)
s <- summarize(group_by(data_extract_merged, taxon), mean_density=mean(density, na.rm=TRUE),
                                                     median_density=median(density, na.rm=TRUE),
                                                     maximum_density=max(density, na.rm=TRUE),
                                                     count_of_200-m_samples_where_present=length(density))

selvar <- read.csv(file.path(savepath, paste(vme, "minus", paste(substr(remove, 1,7), collapse=", "), "selvar.csv")))
selvar <- pull(selvar)

# rcl <- matrix(data = c(0,th,th,max(values(distribution), na.rm =TRUE),1,2), nrow=2)
# rec1 <- rast(raster::reclassify(raster(distribution), rcl, include.lowest = TRUE))
# rcl <- matrix(data = c(0,thr,thr,max(values(error), na.rm =TRUE),1,2), nrow=2)
# rec2 <- rast(raster::reclassify(raster(error), rcl, include.lowest = TRUE))
# combo <- terra::ifel(rec1 == 2 & rec2 == 1, 3, 
#        terra::ifel(rec1 == 2 & rec2 == 2 , 2, 1))
# 
# resp_spat <- readOGR(dsn=savepath , 
#   layer=paste(vme, "minus", paste(substr(remove, 1,7), collapse=", "), "resp_spat"))
# asses <- data.frame(raster::extract(as.raster(combo),resp_spat, sp=TRUE))
# colnames(asses)[5]<-"Zone"
# asses1 <- asses %>% group_by(Zone) %>%
#   dplyr::summarise(mean=mean(vmeind_density), stdev=sd(vmeind_density), count=length(vmeind_density))

```

---
title: "Metadata: `r vme`"
---

# Description
This dataset shows the expected distribution of the Vulnerable Marine Ecosystem (VME) known as
```{r echo=FALSE, message=FALSE, warning=FALSE}
vme
```
Other names: Umbellula stands

We followed the approach described in @buhl2019vulnerable and @burgos2020predicting. This VME is defined by occurrence of the following indicators:
```{r echo=FALSE, message=FALSE, warning=FALSE}
indicators
```

# Goal and usage
The goal of this dataset is to highlight areas which may require special management attention on the grounds of a possible high density of VB indicators, i.e., the potential presence of a VME *sensu* @bacoTaylor2023, or similar conservation relevant habitat. The dataset per VB consists of:

-	A raster layer showing the **predicted cumulative density of indicator taxa**, which can be used to identify both general areas of suitable habitat for indicator taxa and to highlight denser hotspots which may be of management interest.

-	A second raster layer is also supplied showing **predictive error**, with areas showing higher values representing areas where the model is less certain about the values it is predicting. This layer can be used to locate areas where additional sampling is recommended to increase certainty but also may highlight areas that are more heterogenous, where more variability should be expected.

We advise that users make use of both layers to inspect their areas of interest. Please contact us if you would like advice or guidance on usage or alternative visualisations.


# Owner and contact information
Havforskningsinstituttet
Fagekspert: Genoveva Gonzalez Mirelis: <genoveva@hi.no>
Datateknisk: Kjell Bakkeplass: <kjell.bakkeplass@hi.no>

# Resolution/Scale
The native resolution of the dataset in raster format is 200 m.
```{r echo=FALSE, message=FALSE, warning=FALSE}
#mna<-200*200
#paste("1:", sprintf("%.0f",mna*2000))
```

# Coverage
The coverage of this dataset is the MAREANO-multibeamed area as of 2020.

The available number of video lines for this area is 2228, which amounts to 6031 200-m video samples.

# Sources and method
For modelling we used Quantile Regression Forests (QRF) as applied in @hengl2018. While regular Random Forests only give a prediction for the mean value of a target variable, QRF allows you to generate predictions at any desired quantile of a conditional distribution of values for the target variable, thus offering increased flexibility and easy error quantification. A model was trained using observed values of VB indicator density extracted from video, and various environmental layers (including representations of geology, oceanography, and bathymetry) as predictors.

Training data was extracted from the quality controlled MarVid (Mareano Video) database. The density data for each taxon for this VB is summarized in the following table. Note that all taxa densities were summed together prior to modelling such that the model is unaware of species (taxon) identity.

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(s)
```


Concerning predictors, we tested the complete set of 122 predictor variables which were used in the biotope model published in 2022 ^[https://mareano.no/en/topics/habitats/general-biotope-map/mareano-general-biotopes--2022-update-technical-summary/predicted-distribution-of-general-biotopes-raster-map]. The final model retained 98 variables. The number of trees was 500. The model was internally evaluated by using the “Out of Bag” procedure and it resulted in the following R^2 (proportion of variance explained) value:
```{r echo=FALSE, message=FALSE, warning=FALSE}
m$r.squared
```
This model was used to generate a median prediction of median density of all relevant VB indicators, at every cell of a grid covering the entire area; that is, a prediction at quantile 0.5. In addition, we extracted the prediction at quantiles 0.159 and 0.841, which give the upper and lower bounds of a 68.27% prediction interval, and according to the “empirical rule” contain all observations that are within one standard deviation of the mean. These bounds were therefore used to compute the error of the predictions.


```{r qrf-sic97-maps, echo=FALSE, fig.width=9, fig.cap="Predictions and prediction interval, or error of density of all VME indicators."}
leg=c("#0000ff", "#0028d7", "#0050af", "#007986", "#00a15e", "#00ca35", 
        "#00f20d", "#1aff00", "#43ff00", "#6bff00", "#94ff00", "#bcff00", 
        "#e5ff00", "#fff200", "#ffca00", "#ffa100", "#ff7900", "#ff5000", "#ff2800", "#ff0000")
library(sp)
par(mfrow=c(1,2), oma=c(0,0,0,0.5), mar=c(0,0,1.5,1))
plot(distribution, col=leg, 
     main="Density prediction", axes=FALSE, box=FALSE)
#points(resp_spat, pch="+")
plot(error, col=rev(bpy.colors()), 
     main="Density prediction error", axes=FALSE, box=FALSE)
#points(resp_spat, pch="+")
```



```{r}
#colnames(asses1) <-c("Zone", "Mean density", "Standard deviation of density", "No. subsamples")
#kable(asses1)
```


# Updating
The dataset will be updated according to the MAREANO Activity Plan.

# Deliverable
## Format
GeoTiff

## Projection
WGS84 UTM 33N

## Restrictions
License

# Links
The workflow can be downloaded/cloned from the following github repository:
https://github.com/GeOnoveva/VME_models.git

Information about environmental predictors can be found here:
https://mareano.no/en/topics/habitats/general-biotope-map/mareano-general-biotopes--2022-update-technical-summary/predicted-distribution-of-general-biotopes-raster-map


# References
