---
author: "Havforskningsinstituttet"
date: "`r format(Sys.Date())`"
output: word_document
bibliography: references.bib
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
# libraries
library(ranger)
library(spm)
library(yardstick)
library(readxl)
library(dplyr)
library(tinytex)
library(knitr)
library(Ckmeans.1d.dp)
library(terra)

# paths
Data <- "G:/R_projects/VH_MarVid/Data"
outpath <- "G:/R_projects/VH_MarVid/Results"
savepath <- "G:/R_projects/VH_MarVid/Intermediate_objects"

# quantiles
quantiles = c((1-.682)/2, 0.5, 1-(1-.682)/2)

# taxonary
taxonary <- read_xlsx(file.path(Data,"Taxonary.xlsx"), sheet = 1) # needs to be closed

# vme and indicators
vme <- "Soft bottom sponge aggregations (boreal ostur)"
remove="nothing"
#remove <- c("Mycale lingua", "Phakellia sp.", "Axinellidae", "Axinella infundibuliformis")
indicators <- taxonary$Reference_List[which(taxonary$VME_BuhlMortensen_etal_2023==vme)]
#indicators <- indicators[-which(indicators%in%remove)]

# model
name <-"m2"
m <- readRDS(file = file.path(outpath, paste(paste(vme,"minus", paste(substr(remove, 1,7), collapse=", ")),
                                             paste0(name, ".rda"))))

# re-generate v_test and v_train
v <- read.csv(file.path(savepath, paste(vme, "minus", paste(substr(remove, 1,7), collapse=", "), "v.csv", collapse=" ")))
v <- v %>% 
      mutate_at(vars(  landscape, sedclass, seddan, sedmil    ), factor)
#data_split <- read.csv(file.path(outpath, paste(vme,"minus", paste(substr(remove, 1,7), collapse=", "), "data split.csv")))
#v_train <- v %>% left_join(filter(data_split, data_split$set=="train")) %>% filter(!is.na(set))
#v_test <- v %>% left_join(filter(data_split, data_split$set=="test")) %>% filter(!is.na(set))

# predict with m3
# vme.qrf2 <- predict(m,  v_test,
# type="quantiles", quantiles=quantiles)$predictions
# v_test$vme.qrf2pr = vme.qrf2[,2]
# v_test$vme.qrf2var = (vme.qrf2[,3]-vme.qrf2[,1])/2
# 
# # test m3
# predicted <- v_test$vme.qrf2pr
# observed <- v_test %>%
# dplyr::select(vmeind_density)
# vecv <- vecv(pull(observed),pull(data.frame(unlist(predicted))))
# ccc <- ccc_vec(pull(observed),predicted)

# threshold
distribution <- rast(file.path(file.path(outpath, name), paste(vme,"minus", paste(remove, collapse=", "), "pred.tif")))
error <- rast(file.path(file.path(outpath, name), paste(vme,"minus", paste(remove, collapse=", "), "pred var.tif")))
val <- values(distribution, na.rm=TRUE) %>% data.frame %>% pull
th<-Ckmeans.1d.dp(val, k=2)$centers[2]
val <- values(error, na.rm=TRUE) %>% data.frame %>% pull
thr<-Ckmeans.1d.dp(val, k=2)$centers[2]

# data summary
sp<-indicators
e <- read.csv(file.path(savepath, "pred_data/e.csv"))
sample_info <- read.csv(file.path(Data,"sample_info.csv"))
sample_info <- sample_info %>% rename(video.line=VL, section=SampID2, mean.latitude=y_coord, mean.longitude=x_coord) %>%
  dplyr::select(video.line, section, mean.latitude, mean.longitude, SampID)
e <- cbind(e,sample_info)
threshold <- vector(mode = "numeric",
                   length = length(sp))
#for (i in 1:length(sp)){
#  threshold[i] <- (data %>% dplyr::filter(taxon==sp[i]) %>% dplyr::pull(density) %>% sort %>% quantile(., c(.01, .05)))[1]
#}
#thresholdt <- data.frame(sp=sp, threshold=as.numeric(threshold))
#data <- read.csv(file.path(Data,"species_densities.csv"))
#data <- data %>% rename(taxon=clean_taxonomy, density = density_n100m2) %>% dplyr::select(SampID, taxon, density)
#data_extract <- data %>% filter(taxon%in%indicators) %>% left_join(e, by = join_by(SampID == section))
#data_extract_merged <-merge(data_extract,thresholdt,by.x="taxon",by.y="sp",all.x=TRUE)
#s <- summarize(group_by(data_extract_merged, taxon), mean=mean(density, na.rm=TRUE),
#                                                     median=median(density, na.rm=TRUE),
#                                                     max=max(density, na.rm=TRUE),
#                                                     count=length(density))

selvar <- read.csv(file.path(savepath, paste(vme, "minus", paste(substr(remove, 1,7), collapse=", "), "selvar.csv")))
selvar <- pull(selvar)
```

---
title: "Produktark: `r vme`"
---

# Description
This dataset shows the expected distribution of the Vulnerable Marine Ecosystem (VME) known as
```{r echo=FALSE, message=FALSE, warning=FALSE}
vme
```

The definition for this VME followed @buhl2019vulnerable and @burgos2020predicting. Therefore, this VME is defined by occurrence of the following indicators:
```{r echo=FALSE, message=FALSE, warning=FALSE}
indicators
```

# Goal and usage
The goal of this dataset is to highlight areas which may require special management attention on the grounds of possible high density of VME indicators, i.e., the potential presence of a VME (@bacoTaylor2023). Areas classified as 3 are deemed to have high density of any of the indicator taxa listed above as well as a narrow prediction interval (high certainty). Areas classified as 2 are deemed to have a high density of any of the indicator taxa as well as a wide prediction interval (lower certainty), areas classified as 1 are deemed to have low density of any of the indicator taxa.

# Owner and contact information
Havforskningsinstituttet
Fagekspert: Genoveva Gonzalez Mirelis: <genoveva@hi.no>
Datateknisk: Kjell Bakkeplass: <kjell.bakkeplass@hi.no>

# Resolution/Scale
The native resolution of the dataset in raster format is 200. The scale of the vectorized version should be as follows:
```{r echo=FALSE, message=FALSE, warning=FALSE}
mna<-8000*8000
paste("1:", sprintf("%.0f",mna*2000))
```

# Coverage
The coverage of this dataset is the entire MAREANO area as of 2020. It is the same as the biotope map which was published in 2022, shown here: (mareano biotoper dekningskart). The available number of video lines for this area is 2228, which amount to 6031 200-m samples.

# Sources and method
This dataset was generated in three main steps:
(1) a model was trained using observed density of VME indicators extracted from video, and various environmental layers (including geology, oceanography, and bathymetry) as predictors.
(2) the model was then used for predicting the density of VME indicators at every cell of a grid covering the entire area, as well as the 67% prediction interval, which is a good estimate for the prediction error
(2) both raster datasets thus generated (predicted density and prediction error) where binned into two categories each, and then combined as follows: high density with low error, high density with high error, and low density irrespective of error
A brief overview is provided here.

## Modelling
Training data was extracted from MarVid, and it can be summarized as follows:
```{r}
#kable(s)
```

For modelling we used Quantile Regression Forests as applied in @hengl2018, which we used to predict the median VME indicator density, as well as the 67% probability upper and lower limits (i.e., the prediction interval).

Predictor variables were the same as were used in the biotope model published in 2022 ^[https://mareano.no/en/topics/habitats/general-biotope-map/mareano-general-biotopes--2022-update-technical-summary/predicted-distribution-of-general-biotopes-raster-map].

The final model included `r length(selvar)` variables. And the number of trees was `r m$num.trees`. The model was internally evaluated by using the "Out of Bag" procedure and it resulted in the following R^2 value:
```{r}
m$r.squared
```


## Visualization of results
The model outputs consist of two floating point, raster layers (continuous variables). Both of these layers were classified into two classes, which can be roughly described as low and high. The threshold between the two classes was calculated using a univariate kmeans algorithm. This algorithm classifies values into k classes (in this case two) of similar means. The center for the high cluster was used as a threshold. The threshold for the median density values was `r th` and the threshold for the error layer was `r thr`, both of which are in units of number of colonies per 100 m2.


# Updating
The dataset will be updated according to the MAREANO Activity Plan.

# Deliverable
## Format
Shapefile (?)

## Projection
WGS84 UTM 33N

## Restrictions
License

## Links
The workflow can be downloaded/cloned from the following github repository:
https://github.com/GeOnoveva/VME_models.git

Information about environmental predictors can be found here:
https://mareano.no/en/topics/habitats/general-biotope-map/mareano-general-biotopes--2022-update-technical-summary/predicted-distribution-of-general-biotopes-raster-map


## References
