---
title: "Metadata (aka Produktark)"
author: "Genoveva Gonzalez Mirelis"
date: "2023-10-27"
output: pdf_document
code_folding: "hide"
---

```{r echo=FALSE}
# libraries
library(ranger)
library(spm)
library(yardstick)
library(readxl)
library(dplyr)
library(tinytex)

# paths
Data <- "G:/R_projects/VH_MarVid/Data"
outpath <- "G:/R_projects/VH_MarVid/Results"
savepath <- "G:/R_projects/VH_MarVid/Intermediate_objects"

# quantiles
quantiles = c((1-.682)/2, 0.5, 1-(1-.682)/2)

# taxonary
taxonary <- read_xlsx(file.path(Data,"Taxonary.xlsx"), sheet = 1) # needs to be closed

# vme and indicators
vme <- "Hard bottom sponge aggregations"
remove <- c("Mycale lingua", "Phakellia sp.", "Axinellidae", "Axinella infundibuliformis")
indicators <- taxonary$Reference_List[which(taxonary$VME_Burgos_etal_2020==vme)]
indicators <- indicators[-which(indicators%in%remove)]

# model
m <- readRDS(file.path(outpath,paste(vme,"minus", paste(substr(remove, 1,7), collapse=", "), "model.rda" )))

# re-generate v_test and v_train
v <- read.csv(file.path(savepath, paste(vme, "minus", paste(substr(remove, 1,7), collapse=", "), "v.csv", collapse=" ")))
v <- v %>% 
      mutate_at(vars(  landscape, sedclass, seddan, sedmil    ), factor)
data_split <- read.csv(file.path(outpath, paste(vme,"minus", paste(substr(remove, 1,7), collapse=", "), "data split.csv")))
v_train <- v %>% left_join(filter(data_split, data_split$set=="train")) %>% filter(!is.na(set))
v_test <- v %>% left_join(filter(data_split, data_split$set=="test")) %>% filter(!is.na(set))

# predict with m3
vme.qrf2 <- predict(m,  v_test,
type="quantiles", quantiles=quantiles)$predictions
v_test$vme.qrf2pr = vme.qrf2[,2]
v_test$vme.qrf2var = (vme.qrf2[,3]-vme.qrf2[,1])/2

# test m3
predicted <- v_test$vme.qrf2pr
observed <- v_test %>%
dplyr::select(vmeind_density)
vecv <- vecv(pull(observed),pull(data.frame(unlist(predicted))))
ccc <- ccc_vec(pull(observed),predicted)
```

# Description
This vector dataset shows the predicted "hot spots" for the Vulnerable Marine Ecosystem referred to as:
```{r}
vme
```
The definition for this VME followed

@book{buhl2019vulnerable,
  title={Vulnerable marine ecosystems (VMEs): Coral and sponge VMEs in Arctic and sub-Arctic waters--Distribution and threats},
  author={Buhl-Mortensen, Lene and Burgos, Julian Mariano and Steingrund, Petur and Buhl-Mortensen, P{\aa}l and {\'O}lafsd{\'o}ttir, Steinunn H and Ragnarsson, Stef{\'a}n {\'A}},
  volume={2019519},
  year={2019},
  publisher={Nordic Council of Ministers}
}

 and also
 
 @article{burgos2020predicting,
  title={Predicting the distribution of indicator taxa of vulnerable marine ecosystems in the Arctic and sub-arctic waters of the Nordic Seas},
  author={Burgos, Julian M and Buhl-Mortensen, Lene and Buhl-Mortensen, P{\aa}l and {\'O}lafsd{\'o}ttir, Steinunn H and Steingrund, Petur and Ragnarsson, Stef{\'a}n {\'A} and Skagseth, {\O}ystein},
  journal={Frontiers in Marine Science},
  volume={7},
  pages={131},
  year={2020},
  publisher={Frontiers Media SA}
}

Therefore, this VME is defined by occurrence of the following indicators:
```{r}
indicators
```

with model
```{r}
m
```

with evaluation
```{r}
vecv
```
and
```{r}
ccc
```

