---
author: "Havforskningsinstituttet"
date: "`r format(Sys.Date())`"
output: word_document
bibliography: references.bib
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
# libraries
library(ranger)
library(spm)
library(yardstick)
library(readxl)
library(dplyr)
library(tinytex)
library(knitr)
library(Ckmeans.1d.dp)
library(terra)

# paths
Data <- "G:/R_projects/VH_MarVid/Data"
outpath <- "G:/R_projects/VH_MarVid/Results"
savepath <- "G:/R_projects/VH_MarVid/Intermediate_objects"

# quantiles
quantiles = c((1-.682)/2, 0.5, 1-(1-.682)/2)

# taxonary
taxonary <- read_xlsx(file.path(Data,"Taxonary.xlsx"), sheet = 1) # needs to be closed

# vme and indicators
vme <- "Soft bottom sponge aggregations (boreal ostur)"
remove="nothing"
#remove <- c("Mycale lingua", "Phakellia sp.", "Axinellidae", "Axinella infundibuliformis")
indicators <- taxonary$Reference_List[which(taxonary$VME_BuhlMortensen_etal_2023==vme)]
#indicators <- indicators[-which(indicators%in%remove)]

# model
name <-"m2"
m <- readRDS(file = file.path(outpath, paste(paste(vme,"minus", paste(substr(remove, 1,7), collapse=", ")),
                                             paste0(name, ".rda"))))

# re-generate v_test and v_train
# v <- read.csv(file.path(savepath, paste(vme, "minus", paste(substr(remove, 1,7), collapse=", "), "v.csv", collapse=" ")))
# v <- v %>% 
#       mutate_at(vars(  landscape, sedclass, seddan, sedmil    ), factor)
#data_split <- read.csv(file.path(outpath, paste(vme,"minus", paste(substr(remove, 1,7), collapse=", "), "data split.csv")))
#v_train <- v %>% left_join(filter(data_split, data_split$set=="train")) %>% filter(!is.na(set))
#v_test <- v %>% left_join(filter(data_split, data_split$set=="test")) %>% filter(!is.na(set))

# predict with m3
# vme.qrf2 <- predict(m,  v_test,
# type="quantiles", quantiles=quantiles)$predictions
# v_test$vme.qrf2pr = vme.qrf2[,2]
# v_test$vme.qrf2var = (vme.qrf2[,3]-vme.qrf2[,1])/2
# 
# # test m3
# predicted <- v_test$vme.qrf2pr
# observed <- v_test %>%
# dplyr::select(vmeind_density)
# vecv <- vecv(pull(observed),pull(data.frame(unlist(predicted))))
# ccc <- ccc_vec(pull(observed),predicted)

# threshold
distribution <- rast(file.path(file.path(outpath, name), paste(vme,"minus", paste(remove, collapse=", "), "pred.tif")))
error <- rast(file.path(file.path(outpath, name), paste(vme,"minus", paste(remove, collapse=", "), "pred var.tif")))
val <- values(distribution, na.rm=TRUE) %>% data.frame %>% pull
th<-Ckmeans.1d.dp(val, k=2)$centers[2]
# th <- Ckmeans.1d.dp(val, k=2) # used when patches too small to see otherwise
# th <- data.frame(val=val, cl=th$cluster)
# th <- th %>% filter(cl==2)
# th <- min(th$val)

val <- values(error, na.rm=TRUE) %>% data.frame %>% pull
thr<-Ckmeans.1d.dp(val, k=2)$centers[2]
# thr <- Ckmeans.1d.dp(val, k=2)
# thr <- data.frame(val=val, cl=thr$cluster)
# thr <- thr %>% filter(cl==2)
# thr <- min(thr$val)


# data summary
sp<-indicators
e <- read.csv(file.path(savepath, "pred_data/e.csv"))
sample_info <- read.csv(file.path(Data,"sample_info.csv"))
sample_info <- sample_info %>% rename(video.line=VL, section=SampID2, mean.latitude=y_coord, mean.longitude=x_coord) %>%
  dplyr::select(video.line, section, mean.latitude, mean.longitude, SampID)
e <- cbind(e,sample_info)
data <- read.csv(file.path(Data,"species_densities.csv"))
data <- data %>% rename(taxon=clean_taxonomy, density = density_n100m2) %>% dplyr::select(SampID, taxon, density)
threshold <- vector(mode = "numeric",
                   length = length(sp))
for (i in 1:length(sp)){
  threshold[i] <-  quantile(sort(dplyr::pull(dplyr::filter(data, taxon==sp[i]))), c(.01, .05))[1]
}
thresholdt <- data.frame(sp=sp, threshold=as.numeric(threshold))
data_extract <- data %>% filter(taxon%in%indicators) %>% left_join(e, by = join_by(SampID == section))
data_extract_merged <-merge(data_extract,thresholdt,by.x="taxon",by.y="sp",all.x=TRUE)
s <- summarize(group_by(data_extract_merged, taxon), mean=mean(density, na.rm=TRUE),
                                                     median=median(density, na.rm=TRUE),
                                                     max=max(density, na.rm=TRUE),
                                                     count=length(density))

selvar <- read.csv(file.path(savepath, paste(vme, "minus", paste(substr(remove, 1,7), collapse=", "), "selvar.csv")))
selvar <- pull(selvar)
```

---
title: "Metadata: `r vme`"
---

# Description
This dataset shows the expected distribution of the Vulnerable Marine Ecosystem (VME) known as
```{r echo=FALSE, message=FALSE, warning=FALSE}
vme
```
Other names: Svampspikelbunn

We followed the approach described in @buhl2019vulnerable and @burgos2020predicting. This VME is defined by occurrence of the following indicators:
```{r echo=FALSE, message=FALSE, warning=FALSE}
indicators
```

# Goal and usage
The goal of this dataset is to highlight areas which may require special management attention on the grounds of possible high density of VME indicators, i.e., the potential presence of a VME (@bacoTaylor2023). Areas classified as 3 are deemed to have high density of any of the indicator taxa with high certainty. This zone warrants strict management action. Areas classified as 2 are deemed to have a high density of any of the indicator taxa but lower certainty so more density variability should be expected. This zone requires the collection of additional data for a more robust assessment. Areas classified as 1 are deemed to have low density of any of the indicator taxa.

# Owner and contact information
Havforskningsinstituttet
Fagekspert: Genoveva Gonzalez Mirelis: <genoveva@hi.no>
Datateknisk: Kjell Bakkeplass: <kjell.bakkeplass@hi.no>

# Resolution/Scale
The native resolution of the dataset in raster format is 200. The scale of the vectorized version should be as follows:
```{r echo=FALSE, message=FALSE, warning=FALSE}
mna<-200*200
paste("1:", sprintf("%.0f",mna*2000))
```

# Coverage
The coverage of this dataset was intended to be the MAREANO-multibeamed area as of 2020 but this VME required light at the bottom as a predictor, whose extent is limited. Therefore the northern limit had to be set at 75&deg;N instead of 80&deg;N.

The available number of video lines for this area is 1860, which amount to 5422 200-m video samples.

# Sources and method
This dataset was generated in two/three main steps:

(1) A model was trained using observed values of density of VME indicators extracted from video, and various environmental layers (including geology, oceanography, and bathymetry) as predictors.

(2) The model was used to generate a prediction of density of all VME indicators at every cell of a grid covering the entire area, as well as the 67% prediction interval, which is known to be an adequate estimate of the prediction error.

(3) Both raster datasets thus generated (namely median predicted density and prediction interval) where binned into two categories each, and then combined to derive three management zones.

A brief overview is provided below.

## Step 1: Model training, and predictions
Training data was extracted from MarVid. The density data for each taxon is summarized in the following table. Note that all taxa were pulled together prior to modelling such that the model is oblivious of species (taxon) identity.
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(s)
```

For modelling we used Quantile Regression Forests as applied in @hengl2018, which we used to predict the median VME indicator density, as well as the 67% probability upper and lower limits (i.e., the prediction interval).

Predictor variables were the same as were used in the biotope model published in 2022 ^[https://mareano.no/en/topics/habitats/general-biotope-map/mareano-general-biotopes--2022-update-technical-summary/predicted-distribution-of-general-biotopes-raster-map].

The final model included `r length(selvar)` variables. And the number of trees was `r m$num.trees`. The model was internally evaluated by using the "Out of Bag" procedure and it resulted in the following R^2 value:
```{r echo=FALSE, message=FALSE, warning=FALSE}
m$r.squared
```

## Step 2: Model results
The model outputs consist of two floating point, raster layers (continuous variables), shown below.

```{r qrf-sic97-maps, echo=FALSE, fig.width=9, fig.cap="Predictions and prediction interval, or error of density of all VME indicators."}
leg=c("#0000ff", "#0028d7", "#0050af", "#007986", "#00a15e", "#00ca35", 
        "#00f20d", "#1aff00", "#43ff00", "#6bff00", "#94ff00", "#bcff00", 
        "#e5ff00", "#fff200", "#ffca00", "#ffa100", "#ff7900", "#ff5000", "#ff2800", "#ff0000")
library(sp)
par(mfrow=c(1,2), oma=c(0,0,0,0.5), mar=c(0,0,1.5,1))
plot(distribution, col=leg, 
     main="Density prediction", axes=FALSE, box=FALSE)
#points(resp_spat, pch="+")
plot(error, col=rev(bpy.colors()), 
     main="Density prediction error", axes=FALSE, box=FALSE)
#points(resp_spat, pch="+")
```

## Step 3: Visualization of predictions
Both of these layers were classified into two classes, roughly low values and high values. The threshold between the two classes was calculated using a univariate kmeans algorithm. This algorithm classifies values into k classes (in this case two) of similar means. The center of the high-values cluster was used as a threshold. The threshold for the median density values was `r th` and the threshold for the error layer was `r thr`, both of which are in units of number of colonies per 100 m2.


# Updating
The dataset will be updated according to the MAREANO Activity Plan.

# Deliverable
## Format
GeoTiff

## Projection
WGS84 UTM 33N

## Restrictions
License

# Links
The workflow can be downloaded/cloned from the following github repository:
https://github.com/GeOnoveva/VME_models.git

Information about environmental predictors can be found here:
https://mareano.no/en/topics/habitats/general-biotope-map/mareano-general-biotopes--2022-update-technical-summary/predicted-distribution-of-general-biotopes-raster-map


# References
